---
title: "ventenata response to clearing and vegetation type & test of seeding vs. unseeded control"
author: "Claire Tortorelli"
date: "August 18, 2021"
output:
  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)

library(here)
library(tidyverse)
library(ggplot2)
library(lme4)
library(stringr)
library(nlme)
library(emmeans)
```

read in data
```{r data}
bio <- read_csv(here("data", "biomass_data_2019_2020.csv"))
```

### Organize data for analysis
```{r format data}

#add col for plot number, plot, and treatment: C = community, uncleared; N = no neighbors, cleared (apologies for the confusing codes)

bio$plotno <- factor(substr(bio$plot_quad,5,6))
bio$plot <- factor(substr(bio$plot_quad,1,6))
bio$trt <- factor(substr(bio$plot_quad,8,9))


#add factor for just seeded ("C." or "N.") or unseeded (no "CC" or "NN")
bio$trt2 <- recode(bio$trt, 'C.'='seeded', 'N.' = 'seeded', 'CC' = 'unseeded', 'NN'='unseeded')

#convert veg type to factor
bio$vegtype <- factor(substr(bio$plot_quad,1,4), levels = c("ARRI", "ARAR" , "SEEP"))

#remove ARAR03_N+1 from analysis - rodent activity overturned the entire quadrat.
bio <- bio[-which(bio$plot_quad == "ARAR03_N.1"),]

#Adjust for zeros using stahel method for log transformation
bio$stahel_other20.bio <- bio$resident20_g + quantile(bio$resident20_g, .25)^2/quantile(bio$resident20_g, .75) #there are 6 zeros, all in cleared subplots

# 21 zeros
bio$stahel_vd19biomass <- bio$vedu19_g + quantile(bio$vedu19_g, .25)^2/quantile(bio$vedu19_g, .75) 


bio$stahel_vd20biomass <- bio$vedu20_g + 0.01
#decided not to use the Stahel method for 2020 vedu biomass because it added 0.3g which was quite a bit more than the other biomass adjustments (closer to 0.01)

#explore data
bio %>% group_by(vegtype, trt) %>%
  summarise(mean = mean(stahel_vd20biomass))
```

### Fit model

Model VEDU biomass response to removal treatment by veg type in seeded subplots
split block design receives random effect for block (plotno) and plot
control for 2019 biomass (indicates potential seed bank and microsite suitability)


```{r}

#model treatment 
mtrt1 <- lme(log(stahel_vd20biomass) ~ trt*vegtype + log(stahel_vd19biomass), random = ~ 1|plotno/plot, data = bio[which(bio$trt2 == "seeded"),])

plot(mtrt1) #residuals look symmetrical 
summary(mtrt1)

library(car)
Anova(mtrt1)


mtrt.emm1 <- emmeans(mtrt1, specs = revpairwise ~ trt|vegtype, type = "response", adjust = "none")

summary(mtrt.emm1) 


```


examine with all plots (seeded and unseeded)
```{r}

#model treatment 
mtrt <- lme(log(stahel_vd20biomass) ~ trt*vegtype + log(stahel_vd19biomass), random = ~ 1|plotno/plot, data = bio)

plot(mtrt) #residuals look symmetrical 
summary(mtrt)

mtrt.emm <- emmeans(mtrt, specs = pairwise ~ trt|vegtype, type = "response", adjust = "none")

summary(mtrt.emm) 
#no effect of clearing or vegetation type
#large difference between seeded and unseeded subplots

Anova(mtrt)
```


Model effect of seed addition vs. unseeded controls, not accounting for clearing
```{r}
#model just seeded vs. unseeded treatment 
mtrt2 <- lme(log(stahel_vd20biomass) ~ trt2*vegtype + log(stahel_vd19biomass), random = ~ 1|plotno/plot, data = bio)

plot(mtrt2) #residuals look good
summary(mtrt2)

mtrt2.emm <- emmeans(mtrt2, specs = pairwise ~ trt2|vegtype, type = "response", adjust = "none")

summary(mtrt2.emm)
#strong effect of seeding compared to unseeded controls 233% increase in vedu biomass

Anova(mtrt2, type = 2)
```

Model naturally occurring ventenata 2019 response to resident biomass
```{r}

m2019 <- lme(log(stahel_vd19biomass) ~ log(resident20_g)*vegtype, random = ~ 1|plotno/plot, data = bio %>% subset(trt == "C." | trt == "CC"))

plot(m2019) #residuals look good
summary(m2019)

library(sjPlot)

anova(m2019)
#strong effect of seeding compared to unseeded controls 233% increase in vedu biomass
```

----------------------------------------------------------------------------------------------------

### Plot results


Extract medians for plotting (effect of removal on vedu biomass for seeded treatment) - covariate 2019 biomass fixed
```{r}

(means_table1 = mtrt.emm1$emmeans %>%
summary(infer = c(TRUE, FALSE) ) %>%
as.data.frame() )

#rename levels for plotitng

means_table1$vegtype  <- factor(means_table1$vegtype, levels = c("ARRI", "ARAR", "SEEP"), labels = c("scab-flat", "low sage-steppe", "wet meadow"))



means_table1$trt  <- factor(means_table1$trt, levels = c("C.", "N.", labels = c("uncleared", "cleared")))

```

Extract medians (back transformed) for plotting - including effect of seeded treatment (with unseeded subplots)
```{r}

(means_table = mtrt.emm$emmeans %>%
summary(infer = c(TRUE, FALSE) ) %>%
as.data.frame() )

#rename levels for plotting

means_table$vegtype  <- factor(means_table$vegtype, levels = c("ARRI", "ARAR", "SEEP"), labels = c("scabland", "low sage-steppe", "wet meadow"))



means_table$trt  <- factor(means_table$trt, levels = c("C.", "N.", "CC", "NN"), labels = c("uncleared + seed", "cleared + seed", "uncleared + unseeded", "cleared + unseeded"))

```

Extract contrasts for plotting
```{r}
contrast_table <- mtrt.emm1$contrasts %>%
      summary(infer = c(TRUE, FALSE) ) %>%
      as.data.frame() 

#rename levels for plotting

contrast_table$vegtype  <- factor(contrast_table$vegtype, levels = c("ARRI", "ARAR", "SEEP"), labels = c("scab-flat", "low sage-steppe", "wet meadow"))

contrast_table$contrast  <-"cleared/uncleared"

#create contast tables for the seeded vs unseeded controls

contrast_table2 <- mtrt2.emm$contrasts %>%
      summary(infer = c(TRUE, FALSE) ) %>%
      as.data.frame() 

#rename levels for plotting

contrast_table2$vegtype  <- factor(contrast_table2$vegtype, levels = c("ARRI", "ARAR", "SEEP"), labels = c("scab-flat", "low sage-steppe", "wet meadow"))

contrast_table2$contrast  <-"seeded/unseeded control"

```

plot raw data!

```{r}
#reorder and rename levels for plotting
bio_reorder <- bio
bio_reorder$trt <- factor(bio_reorder$trt, levels = c("C.", "N.", "CC", "NN"), labels = c("uncleared+seed", "cleared+seed", "uncleared control", "cleared control"))

levels(bio_reorder$vegtype) <- c("scab-flat", "low sage-steppe", "wet meadow")
  
(g_raw <- ggplot(bio_reorder, aes(y = vedu20_g, x = vegtype, fill = trt)) +
# Define stock as group this week as well as set x and y axes
geom_boxplot() + # Add points, dodge by group
theme_bw(base_size = 14) +
labs(y = expression(italic("V. dubia")~" biomass (g)"))+
  scale_fill_manual(values = c("#606C38", "#DDA15E", "#354230", "#BC6C25"))+
    guides(color = FALSE)+
theme(legend.direction = "horizontal", # make legend horiz
legend.position = "bottom", # change legend position
panel.grid.minor = element_blank(),
axis.title.x=element_blank(),# Remove gridlines
panel.grid.major.x = element_blank() ))



```
save plot

```{r}
ggsave("raw_vedu2020biomassData.svg",g_raw, width = 7, height = 4, units = "in" )
```

Plot contrasts (seeded subplots)

```{r fig.height=4.5, fig.width=4.5}
( g_contrast = ggplot(contrast_table, aes(y = ratio, x = vegtype, color = vegtype)) +
# Define stock as group this week as well as set x and y axes
geom_point(position = position_dodge(width = .75), size = 3 ) + # Add points, dodge by group
geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), size = 1,
position = position_dodge(width = .75) ) + # Add errorbars, dodge by group
theme_bw(base_size = 14) +
  geom_hline(yintercept = 1, linetype = 2)+
labs(y = expression("ratio of "~italic("V. dubia")~" biomass (cleared/uncleared)"))+
scale_color_manual(values = c("#A6761D", "#E6AB02", "#666666"))+
    guides(color = FALSE)+
theme(legend.direction = "horizontal", # make legend horiz
legend.position = "bottom", # change legend position
panel.grid.minor = element_blank(),
axis.title.x=element_blank(),# Remove gridlines
panel.grid.major.x = element_blank() ))
```



plot unseeded control vs seeded subplot contrasts
```{r fig.height=4.5, fig.width=4.5}
# plot unseeded controls

( g_contrast_control = ggplot(contrast_table2, aes(y = ratio, x = vegtype, color = vegtype)) +
# Define stock as group this week as well as set x and y axes
geom_point(position = position_dodge(width = .75), size = 3 ) + # Add points, dodge by group
geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), size = 1,
position = position_dodge(width = .75) ) + # Add errorbars, dodge by group
theme_bw(base_size = 14) +
  geom_hline(yintercept = 1, linetype = 2)+
labs(y = expression("ratio of "~italic("V. dubia")~" biomass (seeded/unseeded control)"))+
scale_color_manual(values = c("#A6761D", "#E6AB02", "#666666"))+
    guides(color = FALSE)+
theme(legend.direction = "horizontal", # make legend horiz
legend.position = "bottom", # change legend position
panel.grid.minor = element_blank(),
axis.title.x=element_blank(),# Remove gridlines
panel.grid.major.x = element_blank() ))


```
combine figures with cowplot


```{r}
library(cowplot)

svg(filename = "vedu_contrasts.svg",
    width = 7.5,
    height = 3.5)

cowplot::plot_grid(g_contrast, g_contrast_control, labels = c("(a)", "(b)"))

dev.off()
```

Plot naturally occurring vedu 2019 biomass response to vegtype*resident biomass

```{r}
#create new dataframe for plotting
#subset for just subplots where community is intact
newdat.lme = bio %>% subset(trt == "C." | trt == "CC")


newdat.lme = data.frame(resident20_g = newdat.lme$resident20_g,
                        vegtype= newdat.lme$vegtype)

#get predictions
newdat.lme$predlme = predict(m2019, newdata = newdat.lme, level = 0)

#get CIs
des = model.matrix(formula(m2019)[-2], newdat.lme)

predvar = diag( des %*% vcov(m2019) %*% t(des) )


#add the confidence interval limits to the dataset for plotting.

newdat.lme$lower = with(newdat.lme, predlme - 2*sqrt(predvar) )
newdat.lme$upper = with(newdat.lme, predlme + 2*sqrt(predvar) )
```


plot lme
```{r fig.height=4, fig.width=4}
( g_vedu2019 = ggplot(bio %>% subset(trt == "C." | trt == "CC"), aes(y = log(stahel_vd19biomass), x = log(resident20_g), color = vegtype)) +
     geom_ribbon(data = newdat.lme, aes(y = NULL, 
        ymin = lower, ymax = upper, color = NULL, fill = vegtype),
                 alpha = .15) +
      geom_line(data = newdat.lme, aes(y = predlme), size = .75)+

# Define stock as group this week as well as set x and y axes
theme_bw(base_size = 14) +
labs(y = expression(~italic("V. dubia")~" log biomass (2019)"), x = "resident log biomass")+
scale_color_manual(values = c("#A6761D", "#E6AB02", "#666666"))+
    guides(color = FALSE)+
  scale_fill_manual(values = c("#A6761D", "#E6AB02", "#666666"))+
    guides(color = FALSE)+
theme(legend.direction = "horizontal", # make legend horiz
legend.position = "bottom", # change legend position
panel.grid.minor = element_blank(),
panel.grid.major.x = element_blank() ))

```

```{r}
ggsave("2019VEDUvsResidentbiomass.svg", g_vedu2019, width = 4, height = 4, units = "in")


```
